---
title: "Summary Plots and Statistics for CalCurCEAS 2024 gliders"
# subtitle: 'optional subtitle with added details'
author: "Selene Fregosi"
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry: margin = 1cm
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    code_folding: hide
    # keep_md: yes
editor_options:
  chunk_output_type: console
params:
  useParamsFlag: TRUE # TRUE if using params, false if prompting for files/paths
  # filter: TRUE # TRUE to load and compare an unfiltered and filtered AcousticStudy
  log: TRUE # set to TRUE to generate log file
  survey: sg680_CalCurCEAS_Sep2024 # for report printing/building filenames, e.g., SCOPE_2024 
  path_acSt: Q:/CalCurCEAS_fall_2024/analysis/pamguard/acoustic_studies # path to AcousticStudies, e.g., 'R:/McCullough/PAMpal/Acoustic_Studies' 
  acousticStudy: pam20217a_glider_banter_sg680_CalCurCEAS_Sep2024_acousticStudy.rds # name of AcousticStudy RDS file, e.g., 2024_SCOPE_2404_ch5.rds 
  acousticStudyFiltered: pam20217a_glider_banter_sg680_CalCurCEAS_Sep2024_acousticStudyFiltered.rds #name of filtered AcousticStudy RDS file, optional
  dbFile: Q:/CalCurCEAS_fall_2024/analysis/pamguard/databases/pam20217a_glider_banter_sg680_CalCurCEAS_Sep2024.sqlite3 # full path to PG database file, e.g., 'S:/SCOPE_2024_Recording/Database/PAM20210ae_SCOPE_Recording_20240430.sqlite3' 
  path_binaries: Q:/CalCurCEAS_fall_2024/analysis/pamguard/binaries/pam20217a_glider_banter_sg680_CalCurCEAS_Sep2024 # path to PG binaries folder, e.g., 'S:/SCOPE_2024_Recording/Binaries', note must use drive letter (not //)
  path_out: Q:/CalCurCEAS_fall_2024/analysis/event_reports
  snrThreshold: 15 # SNR threshold for "good" clicks
  refSpec: FALSE
  path_to_refSpec: FALSE # path to reference spectra files, e.g., '//piccrp4nas/grampus/llharp/processingCode/llamp/files/reference_spectra'
  refSpecList: !r c() # use !r to call r then
  refSpecSp: !r c() #left it empty on purpose for the moment
  calFile: FALSE
---

# Survey: `r params$survey`

`r paste0("Summaries generated: ", format(Sys.time(), "%Y %b %d %H:%M %Z"))`


## Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 4,
  fig.asp = 1
)

# write progress to a log file (OPTIONAL)
if (exists("log", where = params) & (params$log == TRUE)) {
  logFile <- file.path(params$path_out, 
                       paste0("processing_log_", params$survey, "_", 
                              format(Sys.time(), "%Y%m%d_%H%M%S"), ".txt"))
  processingStartTime <- Sys.time() 
  writeLines(paste("Processing started:", Sys.time()), logFile)
} else if (!exists("log", where = params)){
  params$log <- FALSE # set to false if it doesn't exist
}

```

Load needed libraries.
```{r library, results = 'hide'}
# install specialty packages from github and load
# install can be slow so skip if not needed
# devtools::install_github('taikisan21/PAMpal') # this can be slow so only do if necessary
# install.packages('crputils', repos=c('https://pifsc-protected-species-division.r-universe.dev',
#                                     'https://cloud.r-project.org'))
library(PAMpal)
library(crputils)

# load standard CRAN packages (assuming already installed)
library(tcltk)
library(kableExtra)
library(ggplot2)
library(RSQLite)
library(tuneR)
# library(here) # not used anymore
```

**crputils** version: `r packageVersion("crputils")`

**PAMpal** version: `r packageVersion("PAMpal")`

Build survey-specific variables.
```{r set-vars, results = 'hide'}
# For generic run, prompt/select all files and paths

# if not using params, prompt for paths/files
if ((!exists("useParamsFlag", where = params)) |
    ((exists("useParamsFlag", where = params) & params$useParamsFlag == FALSE))) {
  # select database file
  dbFile <- tk_choose.files(
    default = "", caption = "Select PAMguard database file",
    filters = matrix(c("sqlite3 files", ".sqlite3", "all files", "*"), 2, 2, 
                     byrow = TRUE))
  # select path to binaries to go with the selected database
  path_binaries <- choose.dir(default = "", caption = "Select PAMguard binaries folder")
  
  # select AcousticStudy RDS file
  acStFile <- tk_choose.files(
    default = "", caption = "Select AcousticStudy detection data file",
    filters = matrix(c("rds files", ".rds", "all files", "*"), 2, 2, byrow = TRUE))
  
  # select filtered AcousticStudy RDS file (if applicable)
  acStFileFilt <- tk_choose.files(
    default = "", caption = "Select filtered
                                    AcousticStudy detection data file",
    filters = matrix(c("rds files", ".rds", "all files", "*"), 2, 2, byrow = TRUE))
  
  # select output folder for event tables and event reports
  path_out <- choose.dir(default = "", caption = "Select output event_reports folder")

  # if using params, assemble and check path/file variables
} else if (params$useParamsFlag == TRUE) {
  ###############################
  # if using params, set variables based on params values

  # could add whatever params you may want to assemble names and paths to loop
  # through drifts or deployments with less changing of the Rmd
  # e.g., define single path_pg and assemble filepaths and paths
  # dbFile <- file.path(params$path_pg, "Databases", params$pgDatabaseFileName)
  # path_binaries <- file.path(params$path_pg, 'Binaries')
  ###############################

  # check for specified AcousticStudy file
  if (exists("acousticStudy", where = params) & exists("path_acSt", where = params)) {
    acStFile <- file.path(params$path_acSt, params$acousticStudy)
    if (!file.exists(acStFile)) {
      acStFile <- tk_choose.files(
        default = "", caption = "Select AcousticStudy detection data file",
        filters = matrix(c("rds files", ".rds", "all files", "*"),
          2, 2,
          byrow = TRUE
        )
      )
    }
  }

  # check for specified filtered AcousticStudy file
  if (exists("acousticStudyFiltered", where = params) & exists("path_acSt", where = params)) {
    acStFiltFile <- file.path(params$path_acSt, params$acousticStudyFiltered)
    if (!file.exists(acStFiltFile)) {
      # select filtered AcousticStudy RDS file (if applicable)
      acStFiltFile <- tk_choose.files(
        default = "", caption = "Select filtered
                                    AcousticStudy detection data file",
        filters = matrix(c("rds files", ".rds", "all files", "*"),
          2, 2,
          byrow = TRUE
        )
      )
    }
  }

  # check for specified dbFile or prompt
  if (exists("dbFile", where = params) & (file.exists(params$dbFile))) {
    dbFile <- params$dbFile
  } else {
    dbFile <- tk_choose.files(
      default = "", caption = "Select PAMguard database file",
      filters = matrix(c("sqlite3 files", ".sqlite3", "all files", "*"),
        2, 2,
        byrow = TRUE
      )
    )
  }

  # check for specified binaries folder
  if (exists("path_binaries", where = params) & (dir.exists(params$path_binaries))) {
    path_binaries <- params$path_binaries
  } else {
    path_binaries <- choose.dir(default = "", caption = "Select PAMguard binaries folder")
  }

  # OPTIONAL check for specified calibration file
  if (exists("calFile", where = params) & (params$calFile != FALSE)) {
    if (file.exists(params$acousticStudy_filtered)) {
      calFile <- params$calFile
    } else {
      calFile <- tk_choose.files(
        default = "", caption = "Select calibration file",
        filters = matrix(c("txt files", ".txt", "all files", "*"),
          2, 2,
          byrow = TRUE
        )
      )
    }
  }

  # OPTIONAL reference spectra -set files and colors (will load later)
  if (exists("refSpec", where = params) & (params$refSpec != FALSE)) {
    # read in or define the path to the reference spectra
    if (exists("path_refSpec", where = params) & (dir.exists(params$path_refSpec))) {
      path_refSpec <- params$path_refSpec
    } else {
      path_refSpec <- choose.dir(default = "", caption = "Select PAMguard binaries folder")
    }
    # assemble the reference spectra paths
    if (!exists("refSpecList", where = params) | !exist("refSpecSp", where = params)) {
      stop("Error: reference spectra list or species not specified in params")
    } else {
      refSpecs <- list()
      refSpecs$path_refSpec <- file.path(params$path_to_refSpec) # pull from YAML
      refSpecs$refSpecList <- params$refSpecList # pull from YAML
      refSpecs$refSpecSp <- params$refSpecSp # pull from YAML
      # reference spectra colors - allows for up to 6 ref specs
      # pastel
      # rsPalette <- c('#66c2a5', '#fc8d62', '#8da0cb', '#e78ac3',
      #                '#a6d854', '#ffd92f')
      # bold
      refSpecs$rsPalette <- c(
        "#1b9e77", "#d95f02", "#7570b3", "#e7298a",
        "#66a61e", "#e6ab02"
      )
    }
  }
}

# more misc set up
# set up datebase driver
sqlite <- dbDriver("SQLite")

# set time zone to UTC
Sys.setenv(TZ = "UTC")
```

#### Load and prep AcousticStudies

Load an existing `acSt` PAMpal AcousticStudy object for event processing. Load both the unfiltered `acSt` and filtered `acStFilt` AcousticStudies to compare total detection counts from each; plots and summary tables will be from filtered data. Optionally, this can be turned off with `params$filter = FALSE`. 

```{r prep-acSt, results = 'hide'}
# load existing acSt AcousticStudy (created with
# workflow_generate_acousticStudies.R)
if (file.exists(acStFile)) {
  acSt <- readRDS(acStFile)
  cat("Loaded", acStFile, "\n")
} else {
  stop("No AcousticStudy 'acSt' file available. Exiting report.", "\n")
}
# update file paths if needed
acSt <- updateFiles(acSt, db = dbFile, bin = path_binaries)

if (file.exists(acStFiltFile)) {
  acStFilt <- readRDS(acStFiltFile)
  cat("Loaded", acStFiltFile, "\n")
} else {
  stop("No AcousticStudy 'acStFilt' file available. Exiting report.", "\n")
  # acStFilt <- acSt # **hack for looking at unfiltered data** use this if testing/no filtered AS exists
}

# update file paths if needed (e.g., data was moved on HD or server)
acStFilt <- updateFiles(acStFilt, db = dbFile, bin = path_binaries)

# summarize how many events
nEvents <- length(names(events(acSt)))
# number of events may change after filtering (all clicks may be filtered out)
nEventsFilt <- length(names(events(acStFilt)))
```

Loaded `r acStFile` and `r acStFiltFile`:

**`r nEvents` events** before filtering and **`r nEventsFilt` events** after initial filtering of clicks to remove noise.


#### Load reference spectra, if specified

Load the reference spectra for plotting, if set with `refSpecList`.
```{r load-refSpec}
# refSpecList is specified in YAML params. Can be one or multiple species
# for single species specify as char string without quotations e.g., refSpecPc
# for multiple species, specify with call to R and c()
#     e.g., !r c('meanSpecGm', 'refSpecPc_LLHARP')

if (exists("refSpec", where = params) & (params$calFile != FALSE)) {
  if (!is.null(refSpecs)) {
    for (rs in refSpecs$refSpecList) {
      refSpecs[[rs]] <- read.csv(file.path(refSpecs$path_to_refSpec, paste0(rs, ".csv")))
    }
  }
} else {
  refSpecs <- NULL
}
```

## Summarize and filter events by false killer whale criteria

Containing summary metrics for each event. Some filtering is done based on the values in this table. Events that do not meet the below criteria are removed: 

- Events contain at least 10 signals in at least one detector (clicks or whistles)
- Median peak frequency of echolocation clicks is < 22 kHz 
- Median begin frequency of whistles is < 8.1 kHz
- 50% of whistles have a begin frequency between 4 and 8 kHz
- Concatenated spectrogram slope *in progress*
- Currently two variations of the peak frequency slope are displayed - one for all clicks and one based just on the 25th to 75th quantile clicks

```{r event-table}
# set up output data frame
evTable <- data.frame(
  survey = character(nEventsFilt),
  eventId = character(nEventsFilt)
) # ,
# species = character(nEvents))

# populate with cruise string and event names
evTable$survey <- paste0(params$survey)
evTable$eventId <- names(PAMpal::events(acStFilt))

# get all click data
allClicks <- getClickData(acStFilt)
# pull click summary stats
evClickSummary <- allClicks %>%
  group_by(eventId) %>%
  summarise(
    numClicks = n(),
    # numClicks_d0 = sum(detectorName == 'Click_Detector_0', na.rm = TRUE),
    numClicks_d1 = sum(detectorName == "Click_Detector_1", na.rm = TRUE),
    numClicks_d2 = sum(detectorName == "Click_Detector_2", na.rm = TRUE),
    numClicks_d3 = sum(detectorName == "Click_Detector_3", na.rm = TRUE),
    numClicks_d4 = sum(detectorName == "Click_Detector_4", na.rm = TRUE),
    numClicks_d5 = sum(detectorName == "Click_Detector_5", na.rm = TRUE),
    click_medPeakFreq_kHz = median(peak, na.rm = TRUE),
    click_p25PeakFreq_kHz = quantile(peak, 0.25, na.rm = TRUE),
    click_p75PeakFreq_kHz = quantile(peak, 0.75, na.rm = TRUE)
  )
# add to output evTable
evTable <- evTable %>%
  left_join(evClickSummary, by = c("eventId" = "eventId"))

# get all whistle data
allWhistles <- getWhistleData(acStFilt)
# pull whistle summary stats
evWhistleSummary <- allWhistles %>%
  group_by(eventId) %>%
  summarise(
    numWhistles = n(),
    # whistle_medCenterFreq_Hz = median(freqCenter, na.rm = TRUE),
    # whistle_medMedFreq_Hz = median(freqMedian, na.rm = TRUE),
    whistle_medBeginFreq_Hz = median(freqBeg, na.rm = TRUE),
    whistle_percBeginFreq4to8kHz = sum(freqBeg >= 4000 & freqBeg <= 8000) / n()
    # whistle_percCenterFreq5to7kHz = sum(freqCenter >= 5000 & freqCenter <= 7000)/n(),
    # whistle_percMedianFreq5to7kHz = sum(freqMedian >= 5000 & freqMedian <= 7000)/n()
  )
# add to output evTable
evTable <- evTable %>%
  left_join(evWhistleSummary, by = c("eventId" = "eventId"))

# add some columns that will be filled in during the event summary loop

evTable$click_lmPeakFreqSlope_Hz <- NA
evTable$click_lmPeakFreqIntercept_Hz <- NA
evTable$click_lmPeakFreqRsqd <- NA

# add column for keeping/removing based on median peak frequency
evTable$keep <- FALSE # set all to false, will mark keepers as TRUE in loop
# mark true where events meet criteria
evTable$keep[which( # evTable$numClicks >= 10 &
  (evTable$numClicks_d1 >= 10 |
    evTable$numClicks_d2 >= 10 |
    evTable$numClicks_d3 >= 10 |
    evTable$numClicks_d4 >= 10 |
    evTable$numClicks_d5 >= 10 |
    evTable$numWhistles >= 10) &
    (is.na(evTable$numClicks) | # if there are clicks check for freq criteria
      evTable$click_medPeakFreq_kHz <= 22) &
    (is.na(evTable$numWhistles) | # if there are whistles, check for freq criteria
      evTable$whistle_medBeginFreq_Hz <= 8100 &
        evTable$whistle_percBeginFreq4to8kHz >= .5)
)] <- TRUE

evTableTrimmed <- evTable[which(evTable$keep == TRUE), ]

# save event table as CSV - saves to path_out in params or chosen
write.csv(evTable, file = file.path(
  params$path_out,
  paste0(
    "eventTable_", params$survey, "_",
    Sys.Date(), ".csv"
  )
))
write.csv(evTableTrimmed, file = file.path(params$path_out,
  paste0(
    "eventTableTrimmed_", params$survey,
    "_", Sys.Date(), ".csv"
  )
))
# the above will be overwritten during the loop below, but this creates placeholders
```

**`r nrow(evTableTrimmed)` events** meet the false killer whale filtering 
criteria of minimum number of signals and frequency characteristics. Only events
that meet the criteria will be plotted below. 


## Event summaries

Loop through each detection event (n = `r nrow(evTableTrimmed)`) create plots 
and a table of summary statistics for click and whistle detections. This process
uses AcousticStudy objects that have detection data for each event, and also 
pulls information from the Pamguard binaries associated with each AcousticStudy. 

Click plots (other than the SNR plot) only show clicks with SNRs >= 
`r params$snrThreshold` dB. 

```{r process-events-true, results = 'asis'}
# set up processing order
eventList <- events(acStFilt)
eventNames <- names(events(acStFilt))
# # possible fkw first, then others
# fkwPoss <- eventNames[evTable$keep]
# other <- eventNames[!evTable$keep]
# eventOrder <- c(fkwPoss, other)
# # for testing
# # eventOrder <- eventOrder[c(1:2, 9:10)]

for (iEvent in seq_along(eventNames)){
  # iEvent <- 1 # for testing
  # ("pagebreak \n")

  # extract this event UID string
  eventUID <- eventNames[iEvent]
  evTableIdx <- which(evTable$eventId == eventUID)
  
  # log and console output for progress tracking
  logMsg <- paste0("[", format(Sys.time(), "%H:%M:%S"), "] iEvent = ", 
                   iEvent, " / ", nEventsFilt, " | eventUID = ", eventUID)
  message(logMsg)  # Print to console
  if (params$log == TRUE) {
    cat(logMsg, file = logFile, append = TRUE, sep = "\n")
  }

  # # if first fkw or first other, add header
  # if (eventUID == fkwPoss[1]) {
  #   cat("\n\n### Possible false killer whale events")
  # } else if (eventUID == other[1]) {
  #   cat("\n\n### Other cetacean events")
  #   cat(
  #     "\n\nThese events did not meet the criteria indicating the need for",
  #     "additional review as possible false killer whale events. "
  #   )
  # }

  # set header for this event and print time
  cat("\n\n#### Event ID:   ", eventUID, "\n")
  cat(
    "Time:   ", format(
      eventList[[eventUID]]@ancillary$grouping$start,
      "%Y-%m-%d %H:%M%Z"
    ), " to ",
    format(eventList[[eventUID]]@ancillary$grouping$end, "%Y-%m-%d %H:%M%Z"),
    "\n"
  )

  ###### summarize clicks ######
  cl <- crputils::er_clickSummary(acStFilt, eventUID)

  cat(
    "\nEvent contains", nrow(getClickData(acSt[[eventUID]])),
    "original clicks,", paste0("**", cl$nClicks), "valid clicks** after",
    "filtering noise.\n"
  )
  cat("\n")

  # to log
  if (params$log == TRUE) {
    cat("   Clicks summarized\n", file = logFile, append = TRUE)
  }
  
  ###### summarize whistles ######
  wl <- crputils::er_whistleSummary(acStFilt, eventUID)

  cat("\nEvent contains", paste0("**", wl$nWhistles), "whistles**.\n")
  cat("\n")

    # to log
  if (params$log == TRUE) {
    cat("   Whistles summarized\n", file = logFile, append = TRUE)
  }
  
  ###### click plots and table ######

  # if any clicks...
  if (cl$nClicks > 0 & ("snr" %in% names(cl))) {
    cat("\n##### Click plots and table\n")
    cat(
      "\n SNR histogram includes all filtered clicks. Other plots contain only",
      "clicks with SNR >=", params$snrThreshold, "dB",
      paste0("(**n = ", cl$nGoodClicks, " clicks**).\n")
    )
    cat("\n")

    # plot histogram of all clicks and SNR >= snrThreshold dB cut off
    crputils::er_plotClickSNRHist(cl, params$snrThreshold)

    # if sufficient good clicks...
    if (cl$nGoodClicks > 0) {
      # plot histogram of click durations - good clicks only
      crputils::er_plotClickDurationHist(cl)

      cat("\n")
      cat("\n")


      # Calculate and plot Concatenated and Mean Spectrum for clicks w/ max SNR > 15dB
      # NB: JLKM uses more exaggerated SNR (>40 dB) for BWs bc actual spectra may
      # be noisy for single clicks
      # reducing wl can give more accurate noise floor but 'smoother' spectrum
      # increasing wl will give more exact spectrum but may exclude too many
      # clicks based on SNR (noise measure will overlap with click output; noise
      # is just measured as same wl from start of binary snippet and binaries
      # binary snippets are v short)
      # Trial and error - 256 works ok for LLHARP data = 1.28 ms
      # NB: JLKM uses 500 for beaked whales with samp rate 288k
      avgSpec <- PAMpal::calculateAverageSpectra(acStFilt,
        evNum = eventUID,
        wl = 256, channel = 1,
        norm = TRUE, noise = TRUE,
        sort = TRUE,
        snr = params$snrThreshold,
        plot = c(TRUE, FALSE)
      )
      # plot concatenated spectrogram - plotted within calculation function

      # plot average spectrum separately bc added median and refSpecs
      if (!is.null(avgSpec)) {
        crputils::er_plotSpectra(avgSpec, cl,
          refSpecs = refSpecs,
          xLims = c(0, 100),
          snrThreshold = params$snrThreshold
        )
      }

      # SKIPPING concatenated slope - not useful for me
      # # plot slope of concatenated spectrogram
      # csAll <- crputils::er_concatenatedSlope(cl)
      # # calculate and plot slope for just the 25-75th quantile of clicks
      # csMed <- crputils::er_concatenatedSlope(cl, quantLimit = c(.25, .75))

      # # add model outputs to exported evTable
      # if (!exists("csAll")) {
      #   evTable$click_lmPeakFreqSlope_Hz[evTableIdx] <- round(coef(csAll)[2] * 1000, 2)
      #   evTable$click_lmPeakFreqIntercept_Hz[evTableIdx] <- round(coef(csAll)[1] * 1000, 2)
      #   evTable$click_lmPeakFreqRsqd[evTableIdx] <- round(coef(csAll)[2] * 1000, 2)
      # 
      #   cat("\n Slope of peak frequencies in concatenated spectrogram: \n")
      #   cat("\n")
      #   cat("\n")
      #   cat("         All clicks -", round(coef(csAll)[2] * 1000, 2), "Hz\n")
      #   cat(
      #     "         Only clicks within 25th to 75th quantile -",
      #     round(coef(csMed)[2] * 1000, 2), "Hz\n"
      #   )
      # }

      # NB: JLKM BW approach has additional plots here:
      # IPI
      # Waveform of strongest click
      # Wigner plot
      # We are not including those here because not really useful for FKW, but if
      # interested in adding back in, see:
      #       https://github.com/jlkeating/PAMGuard_Event_Code

      cat("\n")
      cat("\n")

      # create median stats table for clicks with psuedo-SNR dB TK noise cut off
      cat(
        "\n\n##### Median statistics for", cl$nGoodClicks, "high SNR clicks with",
        "SNR >=", params$snrThreshold, "dB."
      )
      cat("\n")

      cat(knitr::kable(cl$mt,
        format = "html", caption = "", align = "l",
        row.names = FALSE
      ) %>%
        kable_styling(
          bootstrap_options = c("basic", "condensed"),
          full_width = F
        ))
    } else { # no good clicks so no summary plots/table
      cat("\nNo clicks of sufficient SNR to plot or summarize.\n")
    }
  } else {
    cat("\nNo clicks present after filtering for noise.\n")
  }


  ###### whistle plots and table ######

  # if whistles present ...
  if (wl$nWhistles > 0) {
    cat("\n##### Whistle plots and table\n")


    # plot whistle contours
    crputils::er_plotWhistleContours(acStFilt, eventUID, xMax = 1.5, yMax = 20)

    # plot histogram of median frequency
    crputils::er_plotWhistleBeginFreqHist(wl)

    cat("\n")
    cat("\n")

    cat("Median statistics for n = ", wl$nWhistles, "whistles.\n")
    # create median stats table for all whistles
    cat(knitr::kable(wl$mt,
      format = "html", caption = "", align = "l",
      row.names = FALSE
    ) %>%
      kable_styling(
        bootstrap_options = c("basic", "condensed"),
        full_width = F
      ))
  } else {
    cat("\nNo whistles present.\n")
  }

  cat("\n")
  cat("\n\n --- \n\n")
  cat("\n")
  
  # clean up after each event
  # remove event-specific variables
    rm(cl, wl, eventUID, evTableIdx)
  if (exists("avgSpec")) rm(avgSpec)
  if (exists("csAll")) rm(csAll)
  if (exists("csMed")) rm(csMed)
    # run garbage collection to free up memory
    invisible(gc(verbose = FALSE))
    #close out graphic device
    # invisible(dev.off())
}
```

## Save final event tables
```{r event-table-final}
# trim again
evTableTrimmed <- evTable[which(evTable$keep == TRUE), ]

# save event table as CSV - overwrites temp files from event-tables chunk above
# and saves to specified/chosen path_out
write.csv(evTable, file = file.path(path_out, paste0("eventTable_", params$survey, 
                                                     "_", Sys.Date(), ".csv")))
write.csv(evTableTrimmed, file = file.path(params$path_out, 
                                           paste0("eventTableTrimmed_", 
                                                  params$survey, "_",
                                                  Sys.Date(), ".csv")))

# close out the log with elapsed time
if (params$log == TRUE) {
  elapsedTime <- Sys.time() - processingStartTime
  logMsg <- paste0("Processing complete: ", Sys.time(), 
                   " (Elapsed time: ", round(elapsedTime, 2), ")")
  cat(logMsg, file = logFile, append = TRUE, sep = "\n")
}
```
